---
title: "Symbiot to host cell ratios (S/H) in *O. faveolata* colonies exposed to elevated nutrients and immune challenge"

author: "Ana M. Palacio-Castro"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 7
    fig_width: 7
    df_print: paged
    toc: yes
editor_options:
  chunk_output_type: console
includes:
       in_header: myheader.html
---

# 1. Calculate S/H cell ratio 

## General project set-up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r libraries , results="hide"}

# Get all libraries and sources required to run the script
#devtools::install_github("jrcunning/steponeR")
        library(tidyverse )
        library(steponeR)
        library(plyr)
        library(reshape2)
        library(ggthemes)
        library(gridExtra)
        library(lme4)
        library(effects)
        library(lmerTest)
        library(emmeans)
        library(car)
```

## Calculate qPCR RATIOS (Symbiont/Coral) 

Get the raw data for Ofav R.Cunning steponeR function:

* Get list of plate files to read

* Calculate the ratios

* Extract the results
 
```{r CalculateRatios}

  Ofav.plates <- list.files(path="Data", pattern="csv", full.names=T)
  # Ofav.plates

  # Run Stepone function to get target ratios
  
  Ofav.Out <- steponeR(files=Ofav.plates, target.ratios=c("D.Ofav"), 
                       fluor.norm=list(D=0, Ofav=0),
                       copy.number=list(D=3, Ofav=1),
                       ploidy=list(D=1, Ofav=2),
                       extract=list(D=0.813, Ofav=0.982))
  
  # Target ratio results
  Ofav<-Ofav.Out$result

```

## Data cleaning

```{r RemoveNTCandControls}
    # 1. Check and remove NTC wells
    
    ntc <- Ofav[which(Ofav$Sample.Name=="NTC"), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(ntc), ])
    
    # 2. Check and remove + Control wells
    
    Positive <- Ofav[which(Ofav$Sample.Name=="+"), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(Positive), ])

    # 3. Remove data for samples in which only one technical replicate amplified

    Ofav$D.Ofav[which(Ofav$D.reps==1)] <- NA
    colnames(Ofav)[which(colnames(Ofav) %in% c("D.Ofav" ))] <- c("D.SH")  
    Ofav$D.SH[is.na(Ofav$D.SH)] <- 0
    Ofav$tot.SH <- Ofav$D.SH
    
    NoHSratio <- Ofav[which(Ofav$tot.SH==0), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(NoHSratio), ])
    
    # 4. Remove samples with SD>1
    StDe1 <- Ofav[which((Ofav$Ofav.CT.sd>0.8)|(Ofav$D.CT.sd>1.0)), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(StDe1), ])
  
```
  
## Labeling and sample information

```{r GetLabelsAndSampleInfo}

# Get the sample information 
  SampleName<-read.csv("Sample_Plates.csv") 

# Create unique sample ID+FileName to relabel samples  
   Ofav$Sample.Plate<-paste(Ofav$Sample.Name,Ofav$File.Name, sep="_")
   
# Replace Sample.Names in qPCR data 

rownames(SampleName) <- SampleName$Sample.Plate
    Ofav$Treatment<-SampleName[as.character(Ofav$Sample.Plate), "Treatment"]
    Ofav$Replicate<-SampleName[as.character(Ofav$Sample.Plate), "Replicate"]
    Ofav$Core<-SampleName[as.character(Ofav$Sample.Plate), "Core"]
    Ofav$Colony<-SampleName[as.character(Ofav$Sample.Plate), "Colony"]
    Ofav$Treatment2<-SampleName[as.character(Ofav$Sample.Plate), "Treatment2"]
    Ofav$C<-SampleName[as.character(Ofav$Sample.Plate), "C"]
    Ofav$RNASeq<-SampleName[as.character(Ofav$Sample.Plate), "Sequenced"]
         
    Ofav$Log_tot.SH <- log10(Ofav$tot.SH)  # Calculate log10 SH ratio
    Ofav$Log_tot.SH[which(Ofav$tot.SH==0)] <- NA
    
```

## Check the data and export it if want a local backup

```{r}

Ofav$Colony <- factor(Ofav$Colony, labels = c("Colony 1", "Colony 2", "Colony 3", "Colony 4"))
Ofav$Treatment <- factor(Ofav$Treatment, labels = c("C", "N"))

summary(Ofav)

N_Colony_Treatment<-Ofav %>%
         group_by(Colony, Treatment, Treatment2) %>%
         dplyr::summarise(N = n())
N_Colony_Treatment

# write.csv(Ofav, file = 'Outputs/OfavSH.csv', row.names = F )
```

# 2. Stats and graphs

Figure 1 shows non-transformed data. Statistical test used log10 transformed data since raw S/H cell ratios have a skewed distribution

## Figure 1
  
```{r}

TotSH <-ggplot(Ofav, aes(Treatment, tot.SH)) +
        stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2) +
        stat_summary(fun.y=mean, geom="point", size =4, alpha=0.5) + 
        #stat_summary(fun.y=mean, geom="line")  +
        theme_bw()+ ylab ("Symbiont to host (S/H) cell ratio") + 
        xlab("") + ggtitle("a.")+
        theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text.x=element_text(size=rel(1.5)),
            axis.text.y=element_text(size=rel(1.5)),
            axis.title.y= element_text(size=rel(1.5)),
            plot.title= element_text(size=rel(1.5)))


SH_Colony <- ggplot(Ofav, aes(Treatment, tot.SH)) +
        stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2) +
        stat_summary(fun.y=mean, geom="point", size =4, alpha=0.5) + 
        geom_point(shape=21)+
        #stat_summary(fun.y=mean, geom="line")  +
       theme_bw()+ ylab ("") + ggtitle("b.")+ xlab("")+
       facet_wrap(~Colony, scales = "free") +
         theme( panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text.x=element_text(size=rel(1.5)),
            axis.text.y=element_text(size=rel(1.5)),
            axis.title.y= element_text(size=rel(1.5)),
            plot.title= element_text(size=rel(1.5)))

Figure_1<-grid.arrange(TotSH,SH_Colony, nrow=1, widths=c(1/3, 2/3))

# ggsave(file="Outputs/Figure1.png", plot=Figure_1 , width=6.0, height=4, dpi=300)

```


## ANOVAs
    
### One way ANOVA to test the effect of nutrients alone: 

#### Normality

```{r}
shapiro.test(Ofav$Log_tot.SH)
# Not significatly different from a normal distribution

```

#### homoscedasticity

```{r}
var.test(Log_tot.SH~Treatment, data=Ofav) 
# No significat differences between two variances

bartlett.test(Log_tot.SH~Treatment, data=Ofav) 
# No significat differences between two variances

leveneTest(Log_tot.SH~Treatment, data=Ofav) 
# No significat differences between two variances

fligner.test(Log_tot.SH~Treatment, data=Ofav) 
# No significat differences between two variances

```

    
```{r}

OnwWayANOVA<-aov(Log_tot.SH~Treatment, data=Ofav)
        summary(OnwWayANOVA)
        
```

Nutrients significantly increased the SH cell ratio

### Two Way ANOVA to test the effects of Nutrients (Treatment) * TGFB (Treatment.2)

```{r}
    
TwoWayANOVA<-aov(Log_tot.SH ~ Treatment+Treatment2, data=Ofav)
        summary(TwoWayANOVA)
        drop1(TwoWayANOVA,~.,test="F")
        
```

Summary(ANOVA) and drop1(ANOVA) give different p values for the model, but in both cases Treatment (Nutrients) has a significant effect on the S/H cell ratio while Treatment.2 (TGFB) does not, nor the interaction.

### Two Way ANOVA to test the effects of Nutrients (Treatment) + Colony
  
```{r Models}

TwoFactorAOVLog<-lm(Log_tot.SH ~ Treatment * Colony, data=Ofav)
        step(TwoFactorAOVLog) # Interaction is not significant
TwoFactorAOVLog<-lm(Log_tot.SH ~ Treatment + Colony, data=Ofav)
        anova(TwoFactorAOVLog)
        summary(TwoFactorAOVLog)
        
plot(Effect(c("Treatment"), TwoFactorAOVLog), x.var="Treatment", multiline=T)
emmip(TwoFactorAOVLog, ~Treatment|Colony, CIs = TRUE, 
      aes(Colony=factor(Treatment))) + theme_gdocs()

    ## Only treatment
    Treatment.emm<-emmeans(TwoFactorAOVLog, ~Treatment)
    Treatment.emm
    pairs(Treatment.emm)

    ## Treatment and colony
    Treatment_Colony.emm<-emmeans(TwoFactorAOVLog, ~Treatment + Colony)
    #pairs(Treatment_Colony.emm)
    LogTukey<-cld(Treatment_Colony.emm, by=NULL) # compact-letter display
    LogTukey<-LogTukey[order(LogTukey$Colony),]
    LogTukey
    #write.csv(LogTukey, "Outputs/LogTukey.csv", row.names = F)

```
